Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2016-07-06 inst/examples/20160715-slides-Krakow.Rmd
ok

[Intro 1/3]

@@

### Data and method

Data:

- Fennica (Finnish National Bibliography), n-number of documents

- Kungliga (National Library of Sweden's library catalogue), n-number of documents

Method:

@@

- xxx

@@

### Argument and findings

- Finnish knowledge production 1640-1828 can only be understood in Swedish context.

@@

@@

--> Deep roots of the diversity of "Finnishness" clearly reflcted in knowledge production, 1640-1828.

@@

Kartta jossa näkyy Turku osana Ruotsin valtakuntaa tähän (Janilla ideoita manuskriptin puolella, mun mielestä ne on hyviä tästä). Voisi miettiä, että tämä kartta-slide korvaisi sitten tuon alla tulevan "Comparing publishing activity in Swedish University town".

@@

[Big Picture 4/4]
### Rise of the octavo sized book as vehicle of Enlightenment

@@

Turku in Sweden -slide (Janin ideoima tulisi tähän)

@@

### Numbers of publications of main publishers in Turku

```{r publishers3-finland-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=13, out.width="150px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers3-finland-1.png

@@

df <- df %>%	
        select(publication_decade, publisher, paper.consumption.km2)

@@

### Comparing how much paper different publishers used on average in their publications in Turku

```{r publisherpapertitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width="150px", fig.show="hold"}

@@

Gävle (gray) during riksdagar (Kungliga) TÄÄ EI TOIMI NYT OLLENKAAN KUN ON AIKA PIENI TOI NOUSU NÄKYVISSÄ TÄSSÄ. MENEEKÖ OIKEIN?

@@

Tämän kohdalta olisi tärkeä, että mukana myös joku slide siitä 1760-1770 ajasta. Pitää myös miettiä mihin kohtaan esitystä tämä osio istuu.

@@

@@

[Languages 1/3]

@@

```{r language1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width="150px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/language1-1.png
catalogue = "Fennica"
df <- df0
langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)

@@

[Languages 2/3]

### Development of use of different languages in Kungliga Biblioteket's Swedish data

```{r language1-kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width="150px", fig.show="hold"}
catalogue = "Kungliga"
  df <- dfl %>% filter(catalog == catalogue)
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

[Languages 3/3]

### Percentages of use of different languages in Finnish and Swedish data across time

TÄMÄ PITÄÄ PÄIVITTÄÄ NIIN ETTÄ ON SE RANSKA JA VENÄJÄ MUKANA!

```{r language-perc, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width="150px", fig.show="hold"}
for (catalogue in c("Fennica", "Kungliga")) {

@@

langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

@@

}
```
---

@@

- known as an form of literature denoting the advancement of reading in Finland in the nineteenth century. Yet, it has been unclear when does it show up as a relevant category.

```{r topics, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics-2.png
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German", "Other")

@@

[Religious texts in Finland 2/3]

@@

- Temporary sermons played a crucial role in early modern knowledge production, but it has not been clearly analysed in what language they were in Finland (and Sweden).

```{r topics2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics2-2.png

@@

[Religious texts in Finland 3/3]

@@

langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}

@@

### University curriculum

@@

[University curriculum 1/6]

@@

df <- df0 %>% filter(catalog == catalogue)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

[University curriculum 2/6]

@@

[University curriculum 3/6]

### Printing of philosophy in different languages in Turku

@@

[University curriculum 4/6]

@@

[University curriculum 5/6]

### Printing of economy in different languages in Turku

@@

[University curriculum 6/6]

### Effects of the Enlightenment on the curriculum in Turku

- decline of metaphysics in favour of social and economic theory is clear in Turku
- the Enlightenment literature that people read in Finland was mainly printed in Stockholm and on the continent.

Tähän top-auktorilistat Fennica ja Kungliga.

@@

[Conclusion 1/2]

@@

- Dramatic events regarding the diversity of "Finnishness" take place in the early nineteenth century. Finland becomes part of Russia and Turku is stripped off of its former glory... But the analysis of Finnish knowledge production 1828-2016 is for another day.

@@

df <- fen %>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01)

@@

[Conclusion 2/2]

@@

- Academy of Finland
- University of Helsinki

>>>>

@@

### Data

Data used (1470-1828):

- Fennica, Finnish National Bibliography, 20328 documents

- Kungliga, National Library of Sweden's catalogue, 73165 documents

### Method

@@

- xxx TODO: WHAT DOES LEO WANT TO SAY HERE?

@@

### Context

- Finnish knowledge production 1640-1828 can only be understood in the Swedish context.

@@

### Findings

@@

--> Deep roots of the diversity of "Finnishness" reflcted in knowledge production, 1640-1828.

@@

### TODO Kartta!!!

jossa näkyy Turku osana Ruotsin valtakuntaa

ts. eri kaupungit ja niiden nimikemäärät kartalla. Kartat voisi tehdä, jos mahdollista, vuosille 1650, 1710, 1767 ja 1800 niin näkyy muutos. Tai sitten video jossa näkyy kehitys. Pohjois-Eurooppa karttapohjana.

Toteutuksesta riippuen voidaan miettiä, jos tämä kartta-slide korvaisi tuon alla tulevan "Comparing publishing activity in Swedish University town".

@@

[Big Picture 4/4]

### Rise of the octavo sized book as vehicle of Enlightenment

@@

### TODO Turku in Sweden!!!

Janin ajatus: Kaaviot Fennicasta ja Kungligasta, jossa lasketaan eri kaupunkien nimikemäärien, paperinkulutuksen ja kustantajien suhteellinen osuus vuoteen 1828. Eli samanlainen kaavio kuin on kielistä ”Book printing in Finland” –osiossa. Laaditaan kaaviot Tukholmasta, Uppsalasta, Lundista, Greifswaldista ja Turusta. Voiko myös tehdä erillisen kaavion, jossa Tukholma on poistettu, josta näkee näiden kilpailevien kakkoskaupunkien tilanteen.

@@

### Title count of the output of publishers in Turku

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}

# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers3-finland-1.png

@@

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper.consumption.km2)

@@

### Comparing paper used on average per publication by publishers

```{r publisherpapertitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="150px", fig.show="hold"}

@@

Gävle (gray) during riksdagar (Kungliga) TODO: TÄÄ EI TOIMI NYT OLLENKAAN KUN ON AIKA PIENI TOI NOUSU NÄKYVISSÄ TÄSSÄ. MENEEKÖ OIKEIN? ERI NÄKÖINEN TUOLLA MANUSKRIPTIN PUOLELLA.

@@

### Number of publishers in Stockholm 1760-1770

TODO: Epäonnistuin tässä tuomaan koodin niin että knittaisi oikein. Täten joudun laittamaan kuvalinkin tähän. Harmittaa. Jotkut muut onnistui niin hyvin...

https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/polev1-3.png

@@

TODO: "Other" kielikategorian selvittäminen seuraavissa kuvissa. Onko siellä joku tietty kieli joka näkyy hyvin selvästi mukana analyysissä? Nousee isoksi kategoriaksi. Nyt oon merkannut, että Kreikan kieli (Greek) mukaan moneen kohtaan. Jos kyseessä joku muu kuin Kreikka niin sitten se. Jos taas paljon kombinaatioita pieniä kieliä, niin antaa siinä tapauksessa olla.

@@

[Languages 1/4]
#TODO: TÄSSÄ OLISI HYVÄ JOS v. 1800 julkaisut ei tippuisi nollaan!

@@

```{r language1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="160px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/language1-1.png
catalogue = "Fennica"
df <- df0
langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)

@@

[Languages 2/4]

### Development of use of different languages in Kungliga data

```{r language1-kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="150px", fig.show="hold"}
catalogue = "Kungliga"
  df <- dfl %>% filter(catalog == catalogue)
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

[Languages 3/4]

### Different languages in Fennica and Kungliga (percentage)
TODO: GREEK mukaan (tai jos joku muu iso kieli niin se)!

```{r language-perc, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=12, out.width="150px", fig.show="hold"}
for (catalogue in c("Fennica", "Kungliga")) {

@@

langs <- c("Finnish", "Swedish", "Latin", "German", "French", "Russian", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

@@

###TODO: KIELI% !!!!!!!!!!!!!

Sama prosenttianalyysi kielistä kuin edellisessä slidessä mutta paikkakuntakohtaisesti (sekä Fennican että Kungligan mukaan): Turku, Greifswald, Stockholm, Uppsala, Lund. Rajataan analyysi vuoteen 1828.

[Languages 4/4]

### Topic diversity in different languages in Finnish publications
TODO: TÄMÄ PITÄISI RAJATA LOPPUMAAN VUOTEEN 1828.

```{r topics232, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
df <- filter(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1828)%>% select(publication_year, language, subject_topic)
df <- unique(df)
df <- df %>% group_by(publication_year, language) %>% summarise(n = length(unique(unlist(strsplit(subject_topic, ";")))))

df$language[!df$language %in% langs] <- "Other"
df$language <- factor(df$language, levels = langs)

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point() +
       ggtitle("Number of subject topics")

print(p)
```

@@

- Form of literature denoting the advancement of reading in Finland in the nineteenth century. It has been unclear when does it show up as a relevant category.

```{r topics12122, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics-2.png
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German", "Other")

@@

[Religious texts in Finland 2/3]
#TODO: TÄSTÄ ja seuraavasta TOI "IF FENNICA FULL TIME SPAN" KOMENTO POIS. En saanut koodattua sitä. TODO: GREEK mukaan!

@@

- play an important role in early modern knowledge production. It has not been analysed in what language they were in Finland (and Sweden).

```{r topics882, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics2-2.png

@@

[Religious texts in Finland 3/3]
#TODO: GREEK mukaan!

@@

langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}

@@

### University curriculum in Turku 1640-1828

@@

[University curriculum 1/5]

TODO: OISKO NÄMÄ VIELÄ SELVEMPIÄ VÄRIKKÄINÄ BARPLOTTEINA? Leo, lisäksi säädätkö pois ton risuaitahässäkän näistä? En saanut sitä pois. TODO3: GREEK mukaan!

@@

df <- df0 %>% filter(catalog == catalogue)%>%
  filter(publication_year >= 1640 & publication_year <= 1800)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

[University curriculum 2/5]
TODO: GREEK mukaan!

@@

[University curriculum 3/5]

### Printing of philosophy in different languages in Turku
TODO: GREEK mukaan!

@@

[University curriculum 4/5]
TODO: GREEK mukaan!

@@

[University curriculum 5/5]
TODO: GREEK mukaan!

### Printing of economy and social science in different languages in Turku

@@

### Effects of the Enlightenment in Finland

- decline of metaphysics, rise of social and economic theory 
- the Enlightenment literature read in Finland mainly printed elsewhere

TODO: KUVA PITÄIS RAJATA 1640-1828, EN ONNISTUNUT
```{r authors22, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width="150px", fig.show="hold"}
theme_set(theme_bw(20))
for (catalogue in c("Fennica", "Kungliga")) {
  df <- filter(df0, catalog == catalogue)
  top.authors <- names(top(df, field = "author", n = 10))
  dfs <- df %>% filter(author %in% top.authors) %>%
         group_by(author, publication_decade) %>%
         tally() %>%
         arrange(publication_decade)
  p <- ggplot(dfs, aes(x = publication_decade, y = n, fill = author)) +
       geom_bar(stat = "identity", position = "stack", color = "black") +
       xlab("Publication Decade") +
       ylab("Title Count") +
       scale_fill_grey() +
       guides(fill = guide_legend("Author")) +
       ggtitle(paste("Title count top authors\n", catalogue)) +
       xlim(range(df0$publication_year))
  print(p)
}
```

@@

@@

- Dramatic events regarding the diversity of "Finnishness" take place in the early nineteenth century --- but the analysis of Finnish knowledge production 1828-2016 is for another day.

@@

df <- fen %>%
  filter(publication_year >= 1640 & publication_year <= 1800)%>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01)

@@

### Duplicate publications

TODO: General analysis (any simple picture) of duplicates between Kungliga and Fennica based on Hege's deduplication code.

@@

- Academy of Finland
- University of Helsinki
- National Library of Finland
- National Library of Sweden
- Digitalia / The Regional Council of South Savo
